---
title: "Experiment Walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experiment Walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6)
```

This vignette demonstrates a typical workflow for analyzing eye-tracking data from an experiment using the `eyer` package. We will cover loading data, preprocessing, visualizing, and performing event-based analysis.

## Loading Data

First, we load the `eyer` package and the example data. The `eyer` package comes with example data that we can use.

```{r setup}
library(eyer)
library(ggplot2)

# Define the path to the example data
data_path <- system.file("extdata", package = "eyer")

# Load the data
# We use the "test" prefix as the example files are named test_eyer_...
obj <- load_eyer_data(data_path, "test")
```

For this walkthrough, we want to use a specific set of events that simulate a trial-based experiment. We have a separate events file for this. We will load it and replace the events in our `eyer` object.

```{r load_events}
# Load new events file
events_file <- system.file("extdata", "test_eyer_events_v2.csv", package = "eyer")
new_events <- read.table(events_file, sep = ";", header = TRUE, stringsAsFactors = FALSE)

# Replace events in the object
obj$data$events <- new_events

# Let's inspect the events
head(obj$data$events)
```

## Preprocessing

Eye-tracking data often needs preprocessing. Common steps include downsampling to reduce data size and noise, and recalibrating if there was a drift.

### Downsampling

Let's downsample the data to 10Hz (assuming original is higher, e.g., 100Hz or more).

```{r downsample}
# Downsample to keep every 10th sample
obj <- downsample(obj, n = 10)
```

## Visualization

Visualizing the raw gaze data is crucial for a sanity check.

```{r plot_gaze}
plot_gaze(obj) + 
  geom_eyer_monitor(obj) +
  ggtitle("Raw Gaze Data")
```

We can also visualize specific trials. Let's look at "Trial_1".

```{r plot_trial}
# Filter data for Trial 1
trial_start <- obj$data$events$time[obj$data$events$action == "start_trial" & obj$data$events$name == "Trial_1"]
trial_end <- obj$data$events$time[obj$data$events$action == "end_trial" & obj$data$events$name == "Trial_1"]

# We can use filter_times (if available) or subset manually. 
# Here we will subset for plotting using standard ggplot2 filtering on the dataframe if needed, 
# but eyer might have helpers. Let's stick to base R subsetting for clarity if helpers aren't obvious.

# Actually, let's use the events to filter what we plot.
plot_gaze(obj) +
  geom_eyer_monitor(obj) +
  coord_cartesian(xlim = c(0, 1920), ylim = c(0, 1080)) +
  ggtitle("Gaze Data with Monitor Limits")
```

## Area of Interest (AOI) Analysis

We often want to know if the participant looked at specific regions. Let's define a central square AOI and a top-left AOI.

```{r define_aois}
# Define AOIs
center_aoi <- AreaObject("Center", x = c(760, 1160), y = c(340, 740)) # 400x400 square in center (1920x1080)
top_left_aoi <- AreaObject("TopLeft", x = c(0, 400), y = c(680, 1080))

# Add AOIs to the object
obj <- add_area_column(obj, list(center_aoi, top_left_aoi))

# Check if the area column was added
head(obj$data$fixations)
```

Now we can calculate statistics for these areas.

```{r analyze_areas}
# Analyze fixations in areas
fixation_stats <- analyse_fixation_areas(obj)
print(fixation_stats)

# Analyze gaze points in areas
gaze_stats <- analyse_gaze_areas(obj)
print(gaze_stats)
```

## Event-based Analysis

We can analyze data relative to specific events, like stimulus onset.

```{r event_analysis}
# Let's calculate the number of fixations in the Center AOI during each image presentation.

# Get image onset and offset times
img_onsets <- obj$data$events[obj$data$events$action == "image_onset", ]
img_offsets <- obj$data$events[obj$data$events$action == "image_offset", ]

results <- data.frame()

for(i in seq_len(nrow(img_onsets))) {
  t_start <- img_onsets$time[i]
  t_end <- img_offsets$time[i]
  img_name <- img_onsets$name[i]
  
  # Filter fixations for this trial
  trial_fixations <- obj$data$fixations[obj$data$fixations$time >= t_start & obj$data$fixations$time <= t_end, ]
  
  # Count fixations in Center
  n_center <- sum(trial_fixations$area == "Center", na.rm = TRUE)
  
  results <- rbind(results, data.frame(image = img_name, fixations_in_center = n_center))
}

print(results)
```

This demonstrates a basic workflow for analyzing eye-tracking data with `eyer`.
